Air cargo load and route planning in pickup and delivery operations

### 2024-01-29 latex/eswa.tex
GA and tmax

	\item As the number of nodes is small, we have the possibility to check all possible tours, selecting the one that provides the best value for the objective function. For this reason, a maximum runtime $tmax$\/ per tour is established. In the tests carried out in this work, {\color{blue} $tmax \leq [240s, 1200s, 2400s, 3600s]$}.

	\item At each node, we will use a MIP solver or heuristics to check which method offers the best result within the runtime limit. As there are $K!$ tours, each of them with $K+1$\/ legs, each solution per node will have a runtime limit equal to $tmax/(K+1)!$.

	\item Just for speculation, we will especially test the two tours with the shortest distance, corresponding to the TSP solutions at the nodes considered. In these cases, the runtime limit per node will be $tmax/(K+1)$.

	\item {\color{blue} Lately, the time limit to solve a node will be adjusted to be proportional to the sum of volumes of the candidate items of the node in relation to the sum of volumes of the entire tour. }

@@

	\begin{algorithmic}[1]
			\State {\bf ACLP+RPDP} \,\, {\it in:}~$scenario, surplus, tmax$ \, \, {\it out:}~$answer1, answer2$

@@

			\State Let $\pi_{TSP1}$\/ and $\pi_{TSP2}$ be the shortest tours ({\it cfr.} Table \ref{tab:scenarios})  \label{main:shortest}

@@

			\For {each $method$}  \label{main:method}

				\State $f_1 \gets SolveTour(\pi_{TSP1}, L, M, C, N, method, tmax)$  \label{main:f1}

				\State $f_2 \gets SolveTour(\pi_{TSP2}, L, M, C, N, method, tmax)$  \label{main:f2}

				\State $answer1[scenario,surplus,method] \gets max( f_1, f_2  )$  \label{main:a1}

@@

					\State $f_{\pi} \gets SolveTour(\pi, L, M, C, N, method, tmax/K!)$ \label{main:fpi}

@@

				\State $answer2[scenario,surplus,method] \gets \max f_{\pi}$ \label{main:a2}
			\EndFor
	\end{algorithmic}

@@

\begin{table}[h]
	\centering
	\caption{Testing scenarios}  \label{tab:scenarios}
	\begin{tabular}{c c c c c}
		\toprule
		{\bf Scenario} & $K$ & $L$ & $\pi_{TSP1}$ & $\pi_{TSP2}$ \\		
		\midrule
		1 & 2    & \{$0,1,2$\}           & 0 1 2 0  & 0 2 1 0 \\
		2 & 3    & \{$0,1,2,3$\}         & 0 1 2 3 0 & 0 3 2 1 0\\
		3 & 4    & \{$0,1,2,3,4$\}       & 0 4 1 2 3 0 & 0 3 2 1 4 0 \\
		4 & 5    & \{$0,1,2,3,4,5$\}     & 0 4 1 2 5 3 0 & 0 3 5 2 1 4 0 \\
		5 & 6    & \{$0,1,2,3,4,5,6$\}   & 0 4 1 2 6 5 3 0 & 0 3 5 6 2 1 4 0 \\
		6 & 15   & \{$0,1,2,3,4,5,6,7,8,9,10,11,12,13,14$\}   & not known & not known \\
		\bottomrule
	\end{tabular}
\end{table}

@@

The best results corresponding to the shortest tours are stored in $answer1$\/ (line \ref{main:a1}), and those obtained by testing all $K!$\/ tours are stored in $answer2$\/ (line \ref{main:a2}).

@@

$SolveTour$\/ is described in Algorithm \ref{alg:tour}, where $\pi$\/ is a permutation of the nodes (excluding the base) that defines the order of visits in this tour, $method$\/ corresponds to a MIP solver or a heuristic for solving the node-by-node problems, and $tmax$\/ is the runtime limit that will be distributed among the $K+1$\/ legs of the tour.

@@

		\For {$k \gets 0$ to $K$} \label{tour:loop1}	
		\State $L_{\pi_k} \gets L - \{\pi_0,\pi_1,\ldots,\pi_k\}$  \label{tour:lk1}	 \Comment{the set of remaining nodes is updated}	
		\State $T_i^{\pi_k} \gets -1$, $1 \leq i \leq m$ \label{tour:-1}             \Comment{the pallet destination is unset}

@@

		\State $G_2 \gets SolveNode(method, \pi_k, G_1, tmax/(K+1))$  \label{tour:SolveNode}

@@

One of the main objectives of this work was to find a quick heuristic that offers a good-quality solution for the node-by-node problem. Taking this into account, we design algorithms based on four known meta-heuristics: {\it Ant Colony Optimization} (ACO) \citep{Dorigo1992,DorigoManiezzoColorni1996}, {\it Noising Method Optimization} (NMO) \citep{CharonHudry1993,CharonHudry2001,Zhan2020}, {\it Tabu Search} (TS) \citep{Glover1986}, and {\it Greedy Randomized Adaptive Search Procedure} (GRASP) \citep{FeoResende1989}. We considered several ideas from the literature \citep{NiarFreville1997,Fidanova2006,Alonso2019, Zhan2020}, and we were careful to use the same data structures and procedures in all implementations to enforce fair results comparison.

@@

{\color{blue}The arguments $threshold_1$ and $threshold_2$ are the volume thresholds for $\eta_1$ and $\eta_2$ calculations, respectively.

The argument $tmax$ is the maximum time for Shims to solve a node. It is proportional to the sum of the candidate items volumes in each node, in relation to the sum of all nodes $tmax$, and the maximum time to solve the tour.

@@

We ran Algorithm \ref{alg:main} considering the five scenarios from Table \ref{tab:scenarios}, three values for  $surplus$\/ from $\{1.2, 1.5, 2.0\}$, four values for $tmax$\/ from $\{240s, 1200s, 2400s, 3600s\}$, and six different $method$\/ for the node-by-node solution: $Gurobi$\/ (see \ref{solver}), ACO, NMO, TS, GRASP, and $Shims$\/ (Algorithm \ref{alg:shims}).

@@

	\centering
	\caption{ \color{blue}Overall results}  \label{tab:mr}
	\footnotesize
	\begin{tabular}{cccc}	
	\toprule
	{\bf Method} & {\bf Best scenarios} & {\bf Worst scenarios} & {\bf Worst run times (min)}  \\
	\midrule
	NMO           & 4             & 5    & {\color{red}60}  \\
	ACO           & 2, 3          & 4, 5 & {\color{red}25}, {\color{red}61}  \\
	GRASP         & 1             & 4, 5 & {\color{red}28}, {\color{red}55}  \\
	TS            & -             & 5    & {\color{red}44}  \\
	\midrule
	{\it Gurobi}  & 1, 2, 3, 4    & 5    & {\color{red}did not solve}  \\
	{\it Shims}   & 1, 2, 3, 4, 5 & -    & {\bf 3.26} \\
	\bottomrule
	\end{tabular}
\normalsize

@@

>>>>

{\color{violet}
	\item As the number of nodes is small, we have the possibility to check all possible tours, selecting the one that provides the best value for the objective function. For this reason, a maximum time per solution ($tmax$) is established. In the tests carried out in this work, these times were in the range [240s, 1200s, 2400s, 3600s].

	\item At each node, we will use a MIP solver or heuristics to check which method offers the best result within the time limit for each node.

	\item The time limit to solve a node will be adjusted to be proportional to the sum of volumes of the candidate items of the node in relation to the sum of volumes of the entire tour.
}

@@

	\begin{algorithmic}[1]
			\State {\bf ACLP+RPDP} \,\, {\it in:}~$scenario, surplus, tmax, tsp, ntours$ \, \, {\it out:}~$answer$

@@

@@

		{\color{violet}
			\If {$tsp$ = True}

				\State $tmax \gets tmax/ntours$ \Comment{max. tour time to solve the number of tours from the TSP heuristic}

			\Else

				\State $tmax \gets tmax/K!$

			\EndIf
		}
			\For {each $method$}  \label{main:method}

@@

					\State $f_{\pi} \gets SolveTour(\pi, L, M, C, N, method, tmax)$ \label{main:fpi}

@@

				\State $answer[scenario,surplus,method] \gets \max f_{\pi}$ \label{main:a2}
			\EndFor
	\end{algorithmic}

@@

\begin{table}[h]
	\centering
	\caption{Testing scenarios}  \label{tab:scenarios}
	\begin{tabular}{c c c }
		\toprule
		{\bf Scenario} & $K$ & $L$  \\		
		\midrule
		1 & 2    & \{$0,1,2$\}            \\
		2 & 3    & \{$0,1,2,3$\}         \\
		3 & 4    & \{$0,1,2,3,4$\}        \\
		4 & 5    & \{$0,1,2,3,4,5$\}      \\
		5 & 6    & \{$0,1,2,3,4,5,6$\}    \\
		6 & 15   & \{$0,1,2,3,4,5,6,7,8,9,10,11,12,13,14$\}    \\
		\bottomrule
	\end{tabular}
\end{table}

@@

The best results obtained by testing all tours are stored in $answer$\/ (line \ref{main:a2}).

@@

$SolveTour$\/ is described in Algorithm \ref{alg:tour}, where $\pi$\/ is a permutation of the nodes (excluding the base) that defines the order of visits in this tour, $method$\/ corresponds to a MIP solver or a heuristic for solving the node-by-node problems, and $tmax$\/ is the {\color{blue}runtime limit of the tour.}

@@

		{\color{violet}

		\State $V_{\pi}  \gets 0$ \label{tour:Vol} \Comment{tour volume}

		\For {$k \gets 0$ to $K$} \label{tour:loop0}

			\State $v_{\pi_k}  \gets \sum_{j=1}^{n_k} v_j$
			\State $V_{k} \gets V_{k} + v_{\pi_k}$

		\For {$k \gets 0$ to $K$} \label{tour:loop1}

		\State $tnode_{\pi_k} = (v_{\pi_k}/V_{\pi}) * tmax$ \Comment{{\color{blue} node time limit calculation}}

		\State $L_{\pi_k} \gets L - \{\pi_0,\pi_1,\ldots,\pi_k\}$  \label{tour:lk1}	 \Comment{the set of remaining nodes is updated}	
		\State $T_i^{\pi_k} \gets -1$, $1 \leq i \leq m$ \label{tour:-1}             \Comment{the pallet destination is unset}

@@

		\State $G_2 \gets SolveNode(method, \pi_k, G_1, tnode_{\pi_k})$  \label{tour:SolveNode}

@@

One of the main objectives of this work was to find a quick heuristic that offers a good-quality solution for the node-by-node problem. Taking this into account, we design algorithms based on four known meta-heuristics: {\it Ant Colony Optimization} (ACO) \citep{Dorigo1992,DorigoManiezzoColorni1996}, {\it Noising Method Optimization} (NMO) \citep{CharonHudry1993,CharonHudry2001,Zhan2020}, {\it Tabu Search} (TS) \citep{Glover1986}, {\it Greedy Randomized Adaptive Search Procedure} (GRASP) \citep{FeoResende1989}, and the Genetic Algorithm (GA) \citep{holland1992adaptation}. We considered several ideas from the literature \citep{NiarFreville1997, Fidanova2006, Alonso2019, Zhan2020, shah2020genetic}, and we were careful to use the same data structures and procedures in all implementations to enforce fair results comparison.

@@

{\color{violet}

	The arguments $threshold_1$ and $threshold_2$ are the volume thresholds for $\eta_1$ and $\eta_2$ calculations, respectively.

The argument $tmax$ is the maximum time for {\it Shims} to solve a node. It is proportional to the sum of the candidate items volumes in each node, in relation to the sum of all nodes volumes, and the maximum time to solve the tour.

@@

We ran Algorithm \ref{alg:main} considering the five scenarios from Table \ref{tab:scenarios}, three values for  $surplus$\/ from $\{1.2, 1.5, 2.0\}$, four values for $tmax$\/ from $\{240s, 1200s, 2400s, 3600s\}$, and seven different methods for the node-by-node solution: $Gurobi$\/ (see \ref{solver}), ACO, NMO, TS, GRASP, GA and $Shims$\/ (Algorithm \ref{alg:shims}).

@@

	\centering
	\caption{ \color{blue}Overall results}  \label{tab:mr}
	\footnotesize
	\begin{tabular}{cccc}	
	\toprule
	{\bf Method} & {\bf Best scenarios} & {\bf Worst scenarios} & {\bf Worst run times (min)}  \\
	\midrule
	NMO           & 4             & 5    & {\color{red}60}  \\
	ACO           & 2, 3          & 4, 5 & {\color{red}25}, {\color{red}61}  \\
	GRASP         & 1             & 4, 5 & {\color{red}28}, {\color{red}55}  \\
	TS            & -             & 5    & {\color{red}44}  \\
	GA            & -             & 1,2,3,4,5  & {\color{red}did not solve}  \\
	\midrule
	{\it Gurobi}  & 1, 2, 3, 4    & 5    & {\color{red}did not solve}  \\
	{\it Shims}   & 1, 2, 3, 4, 5 & -    & {\bf 3.26} \\
	\bottomrule
	\end{tabular}
\normalsize

@@

{\color{violet}

As to the genetic algorithm exclusion from the comparisons tests, we must consider that:
\vspace{-2mm}

\begin{itemize}
	\setlength\itemsep{-0.3em}	
	\item The palletization planning of 200 to 1000 items per node in a set of 18 aircraft pallets may be an inadequate problem to be solved with GA, because large chromosome sizes, like 18,000-dimension arrays, can pose significant challenges for each generation to evaluate individual fitness, which demonstrated to be resource-intensive.
	\item Large chromosomes can slow GA convergence due to the vast search space, requiring more generations to find optimal solutions.
	\item Efficient encoding and decoding of the problem into the chromosome and decoding solutions is crucial for a 18,000-dimension array. 
	\item Also, the design of the fitness function becomes more critical as chromosome size increases, as it needs to guide the evolutionary process without being overly computationally demanding.
\end{itemize}
